### 2011-10-20 doc/manuscript.tex
This is the version that Ilan spent time on rewriting section Results.

% Use the PLoS provided bibtex style
\bibliographystyle{plos}

@@

We conducted a series of experiments on simulated data to check the reliability of estimates obtained by ClonalOrigin. We found that the population parameters obtained in the first phase of ClonalOrigin are rather reliable, with the exception of the average recombinant tract length, which appeared to be over-estimated (this was observed also in \citep{Didelot2010}). Our experiments also show that the recombination intensity used in our association tests and traced in the genome browser tracks are reliably recovered by ClonalOrigin (Figure S\ref{fig:ri1}). See Materials and Methods for a more detailed report of these results. One issue we

 with a special focus on its ability to detect asymmetric patterns of recombination such as the one we found between SPY and SDE.

Numbers of 
recombinant edges from the simulated recombinant trees of Simulation 2 (see
\textbf{Supplementary Information} for the description of Simulation 2) were
compared with the corresponding numbers of recombinant edges that were estimated
from the simulated data set from Simulation 2 (see the solid-line intervals with
open circles in Figure \ref{fig:h3}).  We noted that tree-topology-changing
recombinant edges were more reliably recovered compared with
tree-topology-not-changing ones (Figure \ref{fig:h3}). All of the
tree-topology-not-changing recombinant edge types were over-estimated (Figure
\ref{fig:h3}B). The heat map of Figure \ref{fig:Heatmap-of-recombination} did
not show how confident of the values we could be. Comparison of numbers of
recombinant edges estimated from the real data set and the corresponding numbers
estimated from Simulation 2 showed the interval estimates for the heat map
values.  Dotted-line intervals from the red and blue colored cells in the heat
map were above and below the reference of the horizontal line at the zero on the
Y-axis, respectively (Figure \ref{fig:h3}).

for each recombinant edge type to assess the effect of posterior recombinant
trees samples with respect to that of prior recombinant tree samples. The ten
\textit{a priori} recombinant trees from Simulation 2 would have balanced
recombining gene transfer events between SPY and SDE with respect to the prior
expected number of recombinant edges. The posterior recombinant trees used in
Simulation 3 would have unbalanced recombining gene transfer between the two
species branches as we observed in the real data analysis. For significantly
strong gene flows such as ones from SPY to SDE a solid-line interval with a
closed circle would be different from its count part of the dotted-line
interval.  Figure \ref{fig:h3} showed that the recombining gene flow from SPY to
SDE was stronger enough to be nearer to the horizontal line at the zero on the
Y-axis. However, the recombining gene transfer from SDE to SPY appeared to be
too weak.  This comparison supported the biased directionality of recombining
gene transfer from SPY to SDE.

@@

%=============================================================================
% Table 1. 
\begin{table}[!ht]
\caption{
\bf{Functional categories asscoiated with 
recombination intensity based on topology-altering recombinant edges.}}
\noindent \begin{centering}
\begin{tabular}{cccl}
\hline 
P-Value & Count & GO term & Description\tabularnewline
\hline 
1.731e-06 & 38 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\tabularnewline
2.775e-06 & 33 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\tabularnewline
4.581e-05 & 22 & GO:0006633 & fatty acid biosynthetic process\tabularnewline
0.0001248 & 20 & GO:0008610 & lipid biosynthetic process\tabularnewline
0.0002675 & 45 & GO:0008643 & carbohydrate transport\tabularnewline
0.000513 & 81 & GO:0006412 & translation\tabularnewline
0.0005617 & 13 & GO:0006814 & sodium ion transport\tabularnewline
0.001019 & 40 & GO:0030529 & ribonucleoprotein complex\tabularnewline
0.001203 & 12 & GO:0031402 & sodium ion binding\tabularnewline
0.00206 & 39 & GO:0003735 & structural constituent of ribosome\tabularnewline
0.002115 & 237 & GO:0008152 & metabolic process\tabularnewline
0.00215 & 26 & GO:0019843 & rRNA binding\tabularnewline
0.00287 & 371 & GO:0016740 & transferase activity\tabularnewline
0.003552 & 66 & GO:0005840 & ribosome\tabularnewline
\hline 
\end{tabular}
\par\end{centering}
\label{tab:functional}
\end{table}
%\clearpage{}
%=============================================================================

%=============================================================================
% Table 2. 
\begin{table}[!ht]
\caption{
\bf{Functional categories asscoiated with 
non-recombining horizontal gene transfer.}}
\noindent \begin{centering}
\begin{tabular}{cccl}
\hline 
P-Value & Count & GO term & Description \\
\hline 
7e-61 &  67 & GO:0006313 & transposition, DNA-mediated\\
3e-58 &  59 & GO:0004803 & transposase activity\\
1e-30 &  52 & GO:0015074 & DNA integration\\
1e-28 &  23 & GO:0032196 & transposition\\
1e-13 & 527 & GO:0003677 & DNA binding\\
2e-12 &  86 & GO:0006310 & DNA recombination\\
1e-11 & 197 & GO:0003676 & nucleic acid binding\\
3e-09 &  36 & GO:0016987 & sigma factor activity\\
3e-09 &  36 & GO:0006352 & transcription initiation\\
3e-08 & 142 & GO:0043565 & sequence-specific DNA binding\\
\hline 
\end{tabular}
\par\end{centering}
\label{tab:go-events}
\end{table}
\clearpage{}
%=============================================================================

@@

% \section*{Supplementary Tables}
% \clearpage{}

%=============================================================================
% Table S1. List of six genomes.
\begin{table}
\caption{
\bf{The six complete bacterial genomes.}
Each bacteria and its genome is referred to as the name at the first column.}
\noindent \begin{centering}
\begin{tabular}{cccc}
\hline 
Name & NCBI accession & Size (base pairs) & Reference\tabularnewline
\hline
SDE1 & CP002215 & 2159491 & Suzuki et al. (2011)\tabularnewline
SDE2 & NC\_012891 & 2106340 & Shimomura et al. (2011)\tabularnewline
SDD & N/A$^a$ & 2141837 & Suzuki et al. (2011)\tabularnewline
SPY1 & NC\_004070 & 1900521 & Beres et al. (2002)\tabularnewline
SPY2 & NC\_008024 & 1937111 & Beres et al. (2006)\tabularnewline
SEE & NC\_012471 & 2253793 & Holden et al. (2009)\tabularnewline
\hline
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$ SDD genome was not yet available from NCBI.
\end{flushleft}
\label{tab:genome}
\end{table}
%\clearpage{}
%=============================================================================

%=============================================================================
% Table S2. ClonalFrame Result.
\begin{table}[!ht]
\caption{
\bf{Parameter estimates from \texttt{ClonalFrame}.}
\noindent \centering{}\begin{tabular}{cc}
\hline
Parameter & Estimates (95\% credibility interval) \tabularnewline
\hline
$\theta^a$ & 92871.84 (N/A)\tabularnewline
$\nu^b$ & 0.134 (0.132, 0.135)\tabularnewline
$R^c$ & 9370 (8950, 9780)\tabularnewline
$\delta^d$ & 362 (347, 377)\tabularnewline
$r/m^e$ & 4.66 (4.45, 4.87)\tabularnewline
$\rho/\theta^f$ & 0.101 (0.096, 0.105)\tabularnewline
TMRCA$^g$ & 0.331 (0.324, 0.337)\tabularnewline
$T^h$ & 1.02 (1.01, 1.04)\tabularnewline
\hline
\end{tabular}
\begin{flushleft}
$^a$ $\theta$ is population mutation rate,

@@

$^h$ $T$ is the total branch length of the inferred species tree.}
\end{flushleft}
\label{tab:clonalframe}
\end{table}
%\clearpage{}
%=============================================================================

%=============================================================================
% Table S. ClonalOrigin's 1st stage result.
\begin{table}[!ht]
\caption{
\bf{Three population parameter estimates from the model-based approach.}
The estimates of the three parameters were based on one of two independent
MCMC chains.  We considered the similar 
results of the two independent chains as confirmation of 
the convergence of the first stage of MCMC.}
\noindent \begin{centering}
\begin{tabular}{ccc}
\hline
& Replicate 1 & Replicate 2\tabularnewline
Parameter & Estimate (IQR$^a$) & Estimate (IQR)\tabularnewline
\hline
$\theta^b$ & 0.081 (0.067, 0.094) & 0.081 (0.067, 0.099)\tabularnewline
$\rho^c$ & 0.012 (0.006, 0.019) & 0.012 (0.006, 0.020)\tabularnewline
$\delta^d$ & 744 (346, 2848) & 723 (348, 2870)\tabularnewline
\hline
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$ IQR stands for interquartile range.

@@

%=============================================================================
% Table S3. Gene Tree topologies
\begin{table}
\caption{
\bf{Posterior probability of gene tree topologies.}
See Table S\ref{tab:genome} for the reference of the bacteria
individual names.}
\noindent \begin{centering}
\begin{tabular}{cc}
\hline 
Gene tree topologies & Posterior probability\tabularnewline
\hline
(((SDE1,SDE2),SDD),(SPY1,SPY2)) & 66.7\%\tabularnewline
(((SDE1,SDE2),(SPY1,SPY2)),SDD) & 9.3\%\tabularnewline
((SDE1,SDE2),(SDD,(SPY1,SPY2))) & 3.5\%\tabularnewline
(((SDE1,(SDE2,SDD)),(SPY1,SPY2)) & 1.5\%\tabularnewline
(((SDE1,SDD),SDE2),(SPY1,SPY2)) & 1.3\%\tabularnewline
((SDE1,SDD),(SDE2,(SPY1,SPY2))) & 1.2\%\tabularnewline
((((SDE1,SDE2),SPY1),SPY2),SDD) & 1.0\%\tabularnewline
((SDE2,SDD),(SDE1,(SPY1,SPY2))) & 1.0\%\tabularnewline
\hline
\end{tabular}
\par\end{centering}
\label{tab:Gene-tree-topologies}
\end{table}
\clearpage{}
%=============================================================================

%=============================================================================
% Table S4. The values of the heat map.
\begin{table}[!ht]
\caption{
\bf{Ratios of logarithm base 2 of the observed
recombinant edges to the expected number of recombinant edges.}  Figure
\ref{fig:Heatmap-of-recombination} shows the values graphically using heat map.
NA's are for the branch pairs that are impossible under the model-based
approached. Rows are for donor, and columns for recipient.}
\noindent \centering{}\begin{tabular}{cccccccccc}
\hline
& SDE1 & SDE & SDE2 & SD & SDD & ROOT & SPY1 & SPY & SPY2 \tabularnewline
\hline
SDE1& 0.176  &   NA  &1.170  &    NA & 0.7564  & NA & 0.67 &    NA & 0.663\tabularnewline
SDE & -0.118 & 0.081 &-0.202 &     NA&  0.3395 &  NA& -0.15&  0.045& -0.323\tabularnewline
SDE2& 1.201  &   NA  &0.092  &    NA & 0.6978  & NA & 0.60 &    NA & 0.387\tabularnewline
SD  & -0.629 & 0.022 &-0.661 &0.00670&  0.0078 &  NA& -0.94& -0.299& -1.030\tabularnewline
SDD & -0.317 & 0.371 &-0.197 &     NA&  0.0752 &  NA& -1.06& -0.525& -0.915\tabularnewline
ROOT& -1.745 &-0.698 &-1.786 &0.23039& -0.4854 &  NA& -1.37&  0.497& -1.306\tabularnewline
SPY1& 1.245  &1.883  &1.251  &    NA &-2.1997  & NA & 0.06 &    NA & 1.062\tabularnewline
SPY & 0.089  &0.200  &0.298 &0.00098 &-0.7903  & NA &-0.19 &-0.058 &-0.159\tabularnewline
SPY2& 1.174  &1.938  &1.287  &    NA &-1.9740  & NA & 1.07 &    NA & 0.063\tabularnewline
\hline
\end{tabular}
\label{tab:heatmap}
\end{table}
%\clearpage{}
%=============================================================================

%=============================================================================
% Table S5. The values of the heat map of actual counts.
\begin{table}[!ht]
\caption{
\bf{The observed number of recombinant edges.}
The zeros are for the branch pairs that are impossible under the model-based
approached. Rows are for donor, and columns for recipient.}
\noindent \centering{}\begin{tabular}{cccccccccc}
\hline
& SDE1 & SDE & SDE2 & SD & SDD & ROOT & SPY1 & SPY & SPY2 \tabularnewline
\hline
SDE1&8.896 &  0.00 &17.712  & 0.00 & 12.872   & 0 &12.59  &  0.00  &12.53\tabularnewline
SDE &44.626& 136.49& 42.088 &  0.00& 227.061  &  0& 73.39 &  99.95 & 65.25\tabularnewline
SDE2&18.099&   0.00&  8.393 &  0.00&  12.360  &  0& 11.97 &   0.00 & 10.36\tabularnewline
SD  &9.671 &101.74 & 9.464  &31.26 &115.796   & 0 &13.68  & 98.38  &12.86\tabularnewline
SDD &45.539& 168.58& 49.513 &  0.00& 194.706  &  0& 43.09 &  67.14 & 47.65\tabularnewline
ROOT&43.553& 602.67& 42.337 &765.53& 802.813  &  0& 98.86 &2146.01 &103.60\tabularnewline
SPY1&40.800&  11.46& 40.973 &  0.00&   4.371  &  0& 21.15 &   0.00 & 42.35\tabularnewline
SPY &57.618& 262.08& 66.594 & 30.92& 163.518  &  0& 83.56 & 207.00 & 85.10\tabularnewline
SPY2&38.837&  11.90& 42.000 &  0.00&   5.111  &  0& 42.45 &   0.00 & 21.19\tabularnewline
\hline
\end{tabular}
\label{tab:obsheatmap}
\end{table}
%\clearpage{}
%=============================================================================

%=============================================================================
% Table S7. Mowgli - Recombining Transfer
\begin{table}[!ht]
\caption{
\bf{Functional categories enriched in 
recombining transfer events that are inferred using the parsimony-based
approach.}}
\noindent \begin{centering}
\begin{tabular}{cccl}
\hline 
P-value & Count & GO-term & Description \\
\hline 
7e-09 & 153 & GO:0006412 & translation\\
1e-08 &  67 & GO:0030529 & ribonucleoprotein complex\\
1e-07 &  45 & GO:0019843 & rRNA binding\\
2e-06 &  73 & GO:0003735 & structural constituent of ribosome\\
3e-05 & 157 & GO:0005840 & ribosome\\
7e-05 & 221 & GO:0003723 & RNA binding\\
2e-04 &  11 & GO:0004826 & phenylalanine-tRNA ligase activity\\
3e-04 &  14 & GO:0015413 & nickel-transporting ATPase activity\\
4e-04 &  93 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\\
5e-04 & 104 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\\
\hline 
\end{tabular}
\par\end{centering}
\label{tab:go-events-recombining}
\end{table}
%\clearpage{}
%=============================================================================

%=============================================================================
% Table S. 
\begin{table}[!ht]
\caption{
\bf{Virulence genes inferred to have experienced recombining gene transfer
events.}}
\noindent \begin{centering}
\begin{tabular}{lll}
\hline 
Genes & Gene Accessions\tabularnewline
\hline 
putative calcium transporter (MgtA) & SpyM3\_0440 & SPY1 to SDE1\\
putative dipeptidase & SpyM3\_0465 & SPY to SDE\\
superoxide dismutase & SpyM3\_1071 & SPY to SDE\\
putative internalin A precursor & SpyM3\_1035 & SPY to SDE\\
putative heat shock protein (chaperonin GroEL) & SpyM3\_1765 & SPY to SDE1\\
pyruvate formate lyase & SpyM3\_1749 & No specific directions\\
putative glucose kinase & SpyM3\_1180 & SPY to SDE\\
\hline 
\end{tabular}
\par\end{centering}
\label{tab:virrecomb}
\end{table}
%\clearpage{}
%=============================================================================

%=============================================================================
% Table S. 
\begin{table}[!ht]
\caption{
\bf{Virulence genes inferred to have experienced horizontal gene transfer
events.}}
\noindent \begin{centering}
\begin{tabular}{ll}
\hline 
Genes & Gene Accessions\tabularnewline
\hline 
Streptococcal C5a peptidase & SDEG\_0933\\ 
Putative transposase & SpyM3\_1016, SpyM3\_1053, SpyM3\_1054\\
putative response regulator & SpyM3\_0174\\
putative maltose operon transcriptional repressor & SpyM3\_0982\\ 
putative histidine kinase & SpyM3\_0171\\
putative cell wall hydrolase & SpyM3\_0731, SpyM3\_0922, SpyM3\_1208\\
genome-encoded streptodornase & SDEG\_0541\\
prophage-associated deoxyribonuclease & SDEG\_1103\\
Protein F2 like fibronectin-binding protein (adhesin and invasin) & SpyM3\_0104\\
\hline 
\end{tabular}
\par\end{centering}
\label{tab:virhgt}
\end{table}
%\clearpage{}
%=============================================================================

@@

\bf{Genes and their parts transferred via homologous

@@

%=============================================================================
% Figure S2. Three population parameters.
\begin{figure}
\begin{center}
\subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1\lyxdot out\lyxdot theta}}

>>>>

% Use the PLoS provided bibtex style
% \bibliographystyle{plos}

@@

We conducted a series of experiments on simulated data to check the reliability
of estimates obtained by the model-based approach. We found that the population
parameters obtained in the first phase were reliable, with the exception that
the average recombinant tract length was over-estimated (this was observed also
in \citep{Didelot2010}). Our experiments also showed that the recombination
intensity used in our functional category association tests had fairly good
correlation between the true value and the simulated ones.  (Figure
S\ref{fig:ri1}). See \textbf{Materials and Methods} for a more detailed report
of these results. One thing we specifically wished to test in these experiments
was whether the model-based approach reliably detected deviations of recombinant
edge counts from their expectation, as observed in the heatmap of Figure
\ref{fig:Heatmap-of-recombination}. Our major concern was that these deviations
were not directly accounted for by the probabilistic model of ClonalOrigin, but
rather indirectly observed by summary of the posterior sample of recombinant
trees.

We simulated the model to generate a hundred simulated data sets using a
collection of recombinant trees sampled from the prior distribution defined by
the species tree and three population parameters, as estimated from our genomic
data.  We then repeat the inference procedure of the second stage of
\texttt{ClonalOrigin} on each data set and, as in the data analysis, and counted
for each donor-recepient branch pair the average number of recombinant edges.
By normalizing the counts by the prior expected count we obtained intervals of
base 2 logarithms of the ratio of the observed number of recombinant edges to
the prior expectionation (Figure \ref{fig:h3}). The comparison showed a good
concordance of the estimated counts and the prior expected counts for all
topology-altering edge types and a clear over-estimation for the counts of
topology-preserving edges. Recall that we made sure to ignore these edges in our
analysis. For each edge type, we also plotted in Figure \ref{fig:h3} the
log-ratio of the count estimated for it in our data analysis and the expected
count (the closed circles without interval bars); these are the values making up
the heatmap of Figure \ref{fig:Heatmap-of-recombination}. Comparing the counts
estimated in the data analysis and the counts estimated from the simulated data
allowed to evaluate the significance of the divergence from expectation observed
in the heatmap.  Recall the heatmap indicated an excess of recombinant edges
between the SPY and SDE clades in both directions. Our simulations indicated
that the excess observed in the direction SDE to SPY was much less significant,
as the counts estimated from real data often falled very close to the
corresponding counts estimated from simulated data. The excess observed in the
direction SPY to SDE appeared not only to be larger, but also more significant.
This further supported our claim for the net directionality in recombining gene
transfer between these two species.

We wished to use simulated data to test not only the ability of the ClonalOrigin
sample to \textit{detect} deviations from expected counts of recombinant edges, but
also its ability to reliably \textit{quantify} this deviation. For this purpose we
generated a hundred simulated data sets using a collection of recombinant trees
sampled from the posterior sample of ClonalOrigin on the genomic data. The
difference between the two simulated batches was that in the first one, the
recombinant trees used represent the prior distribution, and in the second one,
they represented the posterior distribution, which we knew deviated from the
prior. Again, we recorded the recombinant edge counts obtained by ClonalOrigin
and normalized these counts by the prior expected counts (intervals with closed
circles in in Figure \ref{fig:h3}). We expected these counts to be in better agreement with
the counts observed in the data analysis. For recombinant edges that showed a
defficiency in our data analysis (e.g., from the SPY clade to SDD), the counts
appeared to fit well with the count estimated from genomic data. However, for
recombinant edges that showed an excess in our data analysis (e.g., from the SPY
clade to the SDE clade), the counts appeared to be significantly influenced by the
prior. We noted that the influence of the prior appeared to be even stronger for
topology-preserving recombinant edges. All things considered, we found that defficiency of
a certain recombinant edge type was reliably estimated whereas excesses were 
typically underestimated. Therefore, we concluded that the actual extent to which
recombining transfers between SPY and SDE \textit{preferred} the direction from SPY to
SDE was possibly somewhat underestimated in our analysis.

@@

%=============================================================================
% Table 1. 
\begin{table}[!ht]
\caption{
{\bf Functional categories associated with 
recombination intensity based on topology-altering recombinant edges.}}
\noindent \begin{centering}
\begin{tabular}{cccl}
\hline 
P-Value & Count & GO term & Description\tabularnewline
\hline 
1.731e-06 & 38 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\tabularnewline
2.775e-06 & 33 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\tabularnewline
4.581e-05 & 22 & GO:0006633 & fatty acid biosynthetic process\tabularnewline
0.0001248 & 20 & GO:0008610 & lipid biosynthetic process\tabularnewline
0.0002675 & 45 & GO:0008643 & carbohydrate transport\tabularnewline
0.000513 & 81 & GO:0006412 & translation\tabularnewline
0.0005617 & 13 & GO:0006814 & sodium ion transport\tabularnewline
0.001019 & 40 & GO:0030529 & ribonucleoprotein complex\tabularnewline
0.001203 & 12 & GO:0031402 & sodium ion binding\tabularnewline
0.00206 & 39 & GO:0003735 & structural constituent of ribosome\tabularnewline
0.002115 & 237 & GO:0008152 & metabolic process\tabularnewline
0.00215 & 26 & GO:0019843 & rRNA binding\tabularnewline
0.00287 & 371 & GO:0016740 & transferase activity\tabularnewline
0.003552 & 66 & GO:0005840 & ribosome\tabularnewline
\hline 
\end{tabular}
\par\end{centering}
\label{tab:functional}
\end{table}
%\clearpage{}
%=============================================================================

%=============================================================================
% Table 2. 
\begin{table}[!ht]
\caption{
{\bf Functional categories associated with 
non-recombining horizontal gene transfer.}}
\noindent \begin{centering}
\begin{tabular}{cccl}
\hline 
P-Value & Count & GO term & Description \\
\hline 
7e-61 &  67 & GO:0006313 & transposition, DNA-mediated\\
3e-58 &  59 & GO:0004803 & transposase activity\\
1e-30 &  52 & GO:0015074 & DNA integration\\
1e-28 &  23 & GO:0032196 & transposition\\
1e-13 & 527 & GO:0003677 & DNA binding\\
2e-12 &  86 & GO:0006310 & DNA recombination\\
1e-11 & 197 & GO:0003676 & nucleic acid binding\\
3e-09 &  36 & GO:0016987 & sigma factor activity\\
3e-09 &  36 & GO:0006352 & transcription initiation\\
3e-08 & 142 & GO:0043565 & sequence-specific DNA binding\\
\hline 
\end{tabular}
\par\end{centering}
\label{tab:go-events}
\end{table}
\clearpage{}
%=============================================================================

@@

\section*{Supplementary Tables}

%=============================================================================
% Table S1. List of six genomes.
\begin{table}[!ht]
\caption{
{\bf The six complete bacterial genomes.}
Each bacteria and its genome is referred to as the name at the first column.}
\noindent \begin{centering}
\begin{tabular}{cccc}
\hline 
Name & NCBI accession & Size (base pairs) & Reference\tabularnewline
\hline
SDE1 & CP002215 & 2159491 & Suzuki et al. (2011)\tabularnewline
SDE2 & NC\_012891 & 2106340 & Shimomura et al. (2011)\tabularnewline
SDD & N/A$^a$ & 2141837 & Suzuki et al. (2011)\tabularnewline
SPY1 & NC\_004070 & 1900521 & Beres et al. (2002)\tabularnewline
SPY2 & NC\_008024 & 1937111 & Beres et al. (2006)\tabularnewline
SEE & NC\_012471 & 2253793 & Holden et al. (2009)\tabularnewline
\hline
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$ SDD genome was not yet available from NCBI.
\end{flushleft}
\label{tab:genome}
\end{table}
%\clearpage{}
%=============================================================================

%=============================================================================
% Table S2. ClonalFrame Result.
\begin{table}[!ht]
\caption{
{\bf Parameter estimates from \texttt{ClonalFrame}.}}
\noindent \centering{}\begin{tabular}{cc}
\hline
Parameter & Estimates (95\% credibility interval) \tabularnewline
\hline
$\theta^a$ & 92871.84 (N/A)\tabularnewline
$\nu^b$ & 0.134 (0.132, 0.135)\tabularnewline
$R^c$ & 9370 (8950, 9780)\tabularnewline
$\delta^d$ & 362 (347, 377)\tabularnewline
$r/m^e$ & 4.66 (4.45, 4.87)\tabularnewline
$\rho/\theta^f$ & 0.101 (0.096, 0.105)\tabularnewline
TMRCA$^g$ & 0.331 (0.324, 0.337)\tabularnewline
$T^h$ & 1.02 (1.01, 1.04)\tabularnewline
\hline
\end{tabular}
\begin{flushleft}
$^a$ $\theta$ is population mutation rate,

@@

$^h$ $T$ is the total branch length of the inferred species tree.
\end{flushleft}
\label{tab:clonalframe}
\end{table}
%\clearpage{}
%=============================================================================

%=============================================================================
% Table S. ClonalOrigin's 1st stage result.
\begin{table}[!ht]
\caption{
{\bf Three population parameter estimates from the model-based approach.}
The estimates of the three parameters were based on one of two independent
MCMC chains.  We considered the similar 
results of the two independent chains as confirmation of 
the convergence of the first stage of MCMC.}
\noindent \begin{centering}
\begin{tabular}{ccc}
\hline
& Replicate 1 & Replicate 2\tabularnewline
Parameter & Estimate (IQR$^a$) & Estimate (IQR)\tabularnewline
\hline
$\theta^b$ & 0.081 (0.067, 0.094) & 0.081 (0.067, 0.099)\tabularnewline
$\rho^c$ & 0.012 (0.006, 0.019) & 0.012 (0.006, 0.020)\tabularnewline
$\delta^d$ & 744 (346, 2848) & 723 (348, 2870)\tabularnewline
\hline
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$ IQR stands for interquartile range.

@@

%=============================================================================
% Table S4. Gene tree topologies
\begin{table}[!ht]
\caption{
{\bf Posterior probability of gene tree topologies.}
See Table S\ref{tab:genome} for the reference of the bacteria
individual names.}
\noindent \begin{centering}
\begin{tabular}{cc}
\hline 
Gene tree topologies & Posterior probability\tabularnewline
\hline
(((SDE1,SDE2),SDD),(SPY1,SPY2)) & 66.7\%\tabularnewline
(((SDE1,SDE2),(SPY1,SPY2)),SDD) & 9.3\%\tabularnewline
((SDE1,SDE2),(SDD,(SPY1,SPY2))) & 3.5\%\tabularnewline
(((SDE1,(SDE2,SDD)),(SPY1,SPY2)) & 1.5\%\tabularnewline
(((SDE1,SDD),SDE2),(SPY1,SPY2)) & 1.3\%\tabularnewline
((SDE1,SDD),(SDE2,(SPY1,SPY2))) & 1.2\%\tabularnewline
((((SDE1,SDE2),SPY1),SPY2),SDD) & 1.0\%\tabularnewline
((SDE2,SDD),(SDE1,(SPY1,SPY2))) & 1.0\%\tabularnewline
\hline
\end{tabular}
\par\end{centering}
\label{tab:Gene-tree-topologies}
\end{table}
%\clearpage{}
%=============================================================================

%=============================================================================
% Table S4. The values of the heat map.
\begin{table}[!ht]
\caption{
{\bf Ratios of logarithm base 2 of the observed
recombinant edges to the expected number of recombinant edges.}  Figure
\ref{fig:Heatmap-of-recombination} shows the values graphically using heat map.
NA's are for the branch pairs that are impossible under the model-based
approached. Rows are for donor, and columns for recipient.}
\noindent \centering{}\begin{tabular}{cccccccccc}
\hline
& SDE1 & SDE & SDE2 & SD & SDD & ROOT & SPY1 & SPY & SPY2 \tabularnewline
\hline
SDE1& 0.176  &   NA  &1.170  &    NA & 0.7564  & NA & 0.67 &    NA & 0.663\tabularnewline
SDE & -0.118 & 0.081 &-0.202 &     NA&  0.3395 &  NA& -0.15&  0.045& -0.323\tabularnewline
SDE2& 1.201  &   NA  &0.092  &    NA & 0.6978  & NA & 0.60 &    NA & 0.387\tabularnewline
SD  & -0.629 & 0.022 &-0.661 &0.00670&  0.0078 &  NA& -0.94& -0.299& -1.030\tabularnewline
SDD & -0.317 & 0.371 &-0.197 &     NA&  0.0752 &  NA& -1.06& -0.525& -0.915\tabularnewline
ROOT& -1.745 &-0.698 &-1.786 &0.23039& -0.4854 &  NA& -1.37&  0.497& -1.306\tabularnewline
SPY1& 1.245  &1.883  &1.251  &    NA &-2.1997  & NA & 0.06 &    NA & 1.062\tabularnewline
SPY & 0.089  &0.200  &0.298 &0.00098 &-0.7903  & NA &-0.19 &-0.058 &-0.159\tabularnewline
SPY2& 1.174  &1.938  &1.287  &    NA &-1.9740  & NA & 1.07 &    NA & 0.063\tabularnewline
\hline
\end{tabular}
\label{tab:heatmap}
\end{table}
%\clearpage{}
%=============================================================================

%=============================================================================
% Table S5. The values of the heat map of actual counts.
\begin{table}[!ht]
\caption{
{\bf The observed number of recombinant edges.}
The zeros are for the branch pairs that are impossible under the model-based
approached. Rows are for donor, and columns for recipient.}
\noindent \centering{}\begin{tabular}{cccccccccc}
\hline
& SDE1 & SDE & SDE2 & SD & SDD & ROOT & SPY1 & SPY & SPY2 \tabularnewline
\hline
SDE1&8.896 &  0.00 &17.712  & 0.00 & 12.872   & 0 &12.59  &  0.00  &12.53\tabularnewline
SDE &44.626& 136.49& 42.088 &  0.00& 227.061  &  0& 73.39 &  99.95 & 65.25\tabularnewline
SDE2&18.099&   0.00&  8.393 &  0.00&  12.360  &  0& 11.97 &   0.00 & 10.36\tabularnewline
SD  &9.671 &101.74 & 9.464  &31.26 &115.796   & 0 &13.68  & 98.38  &12.86\tabularnewline
SDD &45.539& 168.58& 49.513 &  0.00& 194.706  &  0& 43.09 &  67.14 & 47.65\tabularnewline
ROOT&43.553& 602.67& 42.337 &765.53& 802.813  &  0& 98.86 &2146.01 &103.60\tabularnewline
SPY1&40.800&  11.46& 40.973 &  0.00&   4.371  &  0& 21.15 &   0.00 & 42.35\tabularnewline
SPY &57.618& 262.08& 66.594 & 30.92& 163.518  &  0& 83.56 & 207.00 & 85.10\tabularnewline
SPY2&38.837&  11.90& 42.000 &  0.00&   5.111  &  0& 42.45 &   0.00 & 21.19\tabularnewline
\hline
\end{tabular}
\label{tab:obsheatmap}
\end{table}
%\clearpage{}
%=============================================================================

%=============================================================================
% Table S7. Mowgli - Recombining Transfer
\begin{table}[!ht]
\caption{
{\bf Functional categories enriched in 
recombining transfer events that are inferred using the parsimony-based
approach.}}
\noindent \begin{centering}
\begin{tabular}{cccl}
\hline 
P-value & Count & GO-term & Description \\
\hline 
7e-09 & 153 & GO:0006412 & translation\\
1e-08 &  67 & GO:0030529 & ribonucleoprotein complex\\
1e-07 &  45 & GO:0019843 & rRNA binding\\
2e-06 &  73 & GO:0003735 & structural constituent of ribosome\\
3e-05 & 157 & GO:0005840 & ribosome\\
7e-05 & 221 & GO:0003723 & RNA binding\\
2e-04 &  11 & GO:0004826 & phenylalanine-tRNA ligase activity\\
3e-04 &  14 & GO:0015413 & nickel-transporting ATPase activity\\
4e-04 &  93 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\\
5e-04 & 104 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\\
\hline 
\end{tabular}
\par\end{centering}
\label{tab:go-events-recombining}
\end{table}
%\clearpage{}
%=============================================================================

%=============================================================================
% Table S. 
\begin{table}[!ht]
\caption{
{\bf Virulence genes inferred to have experienced recombining gene transfer
events.}}
\noindent \begin{centering}
\begin{tabular}{lll}
\hline 
Genes & Gene Accessions\tabularnewline
\hline 
putative calcium transporter (MgtA) & SpyM3\_0440 & SPY1 to SDE1\\
putative dipeptidase & SpyM3\_0465 & SPY to SDE\\
superoxide dismutase & SpyM3\_1071 & SPY to SDE\\
putative internalin A precursor & SpyM3\_1035 & SPY to SDE\\
putative heat shock protein (chaperonin GroEL) & SpyM3\_1765 & SPY to SDE1\\
pyruvate formate lyase & SpyM3\_1749 & No specific directions\\
putative glucose kinase & SpyM3\_1180 & SPY to SDE\\
\hline 
\end{tabular}
\par\end{centering}
\label{tab:virrecomb}
\end{table}
%\clearpage{}
%=============================================================================

%=============================================================================
% Table S. 
\begin{table}[!ht]
\caption{
{\bf Virulence genes inferred to have experienced horizontal gene transfer
events.}}
\noindent \begin{centering}
\begin{tabular}{ll}
\hline 
Genes & Gene Accessions\tabularnewline
\hline 
Streptococcal C5a peptidase & SDEG\_0933\\ 
Putative transposase & SpyM3\_1016, SpyM3\_1053, SpyM3\_1054\\
putative response regulator & SpyM3\_0174\\
putative maltose operon transcriptional repressor & SpyM3\_0982\\ 
putative histidine kinase & SpyM3\_0171\\
putative cell wall hydrolase & SpyM3\_0731, SpyM3\_0922, SpyM3\_1208\\
genome-encoded streptodornase & SDEG\_0541\\
prophage-associated deoxyribonuclease & SDEG\_1103\\
Protein F2 like fibronectin-binding protein (adhesin and invasin) & SpyM3\_0104\\
\hline 
\end{tabular}
\par\end{centering}
\label{tab:virhgt}
\end{table}
%\clearpage{}
%=============================================================================

@@

{\bf Genes and their parts transferred via homologous

@@

%=============================================================================
% Figure S2. Three population parameters.
\begin{figure}[!ht]
\begin{center}
\subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1\lyxdot out\lyxdot theta}}
